---
title: "Long Term Ice Cover Duration"
author: "Marie Rivers"
date: "8/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(ggExtra)
library(lubridate)
```

```{r}
meteorological_data <- read_csv(here("data", "ntl20_v6.csv"))
ice_data <- read_csv(here("data", "ntl33_v7.csv"), na = "-999")

```
Ice duration across different lakes in teh Madison Lake Area
```{r}
ice_duration_graph <- ggplot(data = ice_data, aes(x = lakeid, y = ice_duration)) +
  geom_jitter(aes(color = lakeid))

ggMarginal(ice_duration_graph, type = "histogram", groupColour = TRUE, fill = "white")
```
Find the mean ice cover duration in Madison Lake Area by year:
the mean is the mean of the three lakes for each year.
```{r}
annual_duration <- ice_data %>% 
  group_by(year4) %>% 
  summarise(mean_duration = mean(ice_duration)) %>% 
  filter(mean_duration != "NA")

annual_duration_graph <- ggplot(data = annual_duration, aes(x = year4, y = mean_duration)) +
  geom_point() +
  labs(caption = "There is wide variability in annual mean ice duration on lakes \nin the of Madison Lake Area, however there is a slightly negative \ntrend to these values")

annual_duration_graph
```
Join the mean winter air temperatures to the mean ice cover duration data 
```{r}
annual_temp <- meteorological_data %>% 
  group_by(year4) %>% 
  summarise(mean_air_temp = mean(ave_air_temp_adjusted))

annual_duration_join <- left_join(annual_duration, annual_temp, by = "year4")
```
Create a scatterplot of mean winter temperatures vs. mean ice duration
```{r}
temp_duration_graph <- ggplot(data = annual_duration_join, aes(x = mean_air_temp, y = mean_duration)) +
  geom_point() +
  labs(caption = "The data does not show a correlation between mean winter air temp and ice duration")
temp_duration_graph
```
Explore thaw dates sice 1970
```{r}
ice_since_1970 <- ice_data %>% 
  filter(year4 >= 1970) %>% 
  mutate(ice_off = lubridate::ymd(ice_off)) %>% 
  mutate(ice_off_day = yday(ice_off))

ice_off_day_graph <- ggplot(data = ice_since_1970, aes(x = year4, y = ice_off_day)) +
  geom_point(aes(color = lakeid)) +
  facet_wrap(~lakeid) +
  labs(caption = "there is no temporal trend in the day of year ice is off lake in the Madison Lake Area")
ice_off_day_graph
```
scale_x_date(date_breaks = "5 years",
               date_labels = "%Y") +
               
# `y` lowercase is 2 didget year, `Y` upper case is a 4 year date
ggplot(data = lizard_obs_count, aes(x = date, y = count)) +
  geom_line() +
  scale_x_date(date_breaks = "3 years",
               date_labels = "%Y")
